---
title: "R Notebook"
output: html_notebook
---

# SimBu tutorial

<http://omnideconv.org/SimBu/articles/simulator_documentation.html>

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("SimBu")
```

```{r}
counts <- Matrix::Matrix(matrix(rpois(3e5, 5), ncol=300), sparse = TRUE)
tpm <- Matrix::Matrix(matrix(rpois(3e5, 5), ncol=300), sparse = TRUE)
tpm <- Matrix::t(1e6*Matrix::t(tpm)/Matrix::colSums(tpm))

colnames(counts) <- paste0("cell_",rep(1:300))
colnames(tpm) <- paste0("cell_",rep(1:300))
rownames(counts) <- paste0("gene_",rep(1:1000))
rownames(tpm) <- paste0("gene_",rep(1:1000))

annotation_dataframe <- data.frame("ID"=paste0("cell_",rep(1:300)), 
                         "cell_type"=c(rep("T cells CD4",50), 
                                       rep("T cells CD8",50),
                                       rep("Macrophages",100),
                                       rep("NK cells",10),
                                       rep("B cells",70),
                                       rep("Monocytes",20)))
```

```{r}
counts
```

```{r}
annotation_dataframe
```

```{r}
library(SimBu)
# use local data to build dataset
dataset <- dataset(annotation = annotation_dataframe, count_matrix = counts, name = "test_dataset")
simulation <- simulate_bulk(data = dataset, scenario = "random", scaling_factor = "NONE")
```

```{r}
simulation
```

## Creating SimBu format of the dataset

```{r}
ds <- SimBu::dataset(annotation = annotation_dataframe,
                     count_matrix = counts,
                     tpm_matrix = tpm,
                     name = "test_dataset")
```

```{r}
simulation <- SimBu::simulate_bulk(data = ds,
                                   scenario = "random", 
                                   scaling_factor = "NONE", 
                                   ncells=100, 
                                   nsamples = 10, 
                                   BPPARAM = BiocParallel::MulticoreParam(workers = 4),  #this will use 4 threads to run the simulation
                                   run_parallel = TRUE)  
```

SimBu can add mRNA bias by using different scaling factors to the simulations using the `scaling_factor` parameter. A detailed explanation can be found in the [\"Scaling factor\"](http://omnideconv.org/SimBu/articles/simulator_scaling_factors.Rmd) vignette.

Currently there are 6 `scenarios` implemented in the package:

-   *even*: this creates samples, where all existing cell-types in the dataset appear in the same proportions. So using a dataset with 3 cell-types, this will simulate samples, where all cell-type fractions are 1/3. In order to still have a slight variation
    between cell type fractions, you can increase the `balance_uniform_mirror_scenario` parameter (default to 0.01). Setting it to 0 will generate simulations with exactly the same cell type fractions.

-   *random*: this scenario will create random cell type fractions using all present types for each sample. The random sampling is based on the uniform distribution.

-   *mirror_db*: this scenario will mirror the exact fractions of cell types which are present in the provided dataset. If it consists of 20% T cells, 30% B cells and 50% NK cells, all simulated samples
    will mirror these fractions. Similar to the uniform scenario, you can add a small variation to these fractions with the `balance_uniform_mirror_scenario` parameter.

-   *weighted*: here you need to set two additional parameters for the [`simulate_bulk()`](http://omnideconv.org/SimBu/reference/simulate_bulk.html) function: `weighted_cell_type` sets the cell-type you want to be over-representing and `weighted_amount` sets the fraction of this cell-type. You could for example use `B-cell` and `0.5` to create samples, where 50% are B-cells and the rest is filled randomly with other cell-types.

-   *pure*: this creates simulations of only one single cell-type. You have to provide the name of this cell-type with the `pure_cell_type` parameter.

-   *custom*: here you are able to create your own set of cell-type fractions. When using this scenario, you additionally need to provide a dataframe in the `custom_scenario_data` parameter, where each row represents one sample (therefore the number of rows need to match the `nsamples` parameter). Each column has to represent one cell-type, which also occurs in the dataset and describes the fraction of this cell-type in a sample. The fractions per sample need to sum up to 1. An example can be seen here:

```{r}
SimBu::plot_simulation(simulation = simulation)
```
